### 2025-02-20 code/eda-code/eda.qmd
Project Work

---
title: "An example exploratory analysis script"
date: "2024-02-07"
output: html_document
---

This Quarto file loads the cleaned data and does some exploring.

I'm only showing it the way where the code is included in the file. 
As described in the `processing_code` materials, I currently prefer the approach of having R code in a separate file and pulling it in.

But I already had this written and haven't yet re-done it that way. Feel free to redo and send a pull request on GitHub :)

Again, it is largely a matter of preference and what makes the most sense to decide if one wants to have code inside Quarto files, or as separate R files.
And sometimes, an R script with enough comments is good enough and one doesn't need a Quarto file.

Also note that while here I split cleaning and exploring, this is iterative. You saw that as part of the processing, we already had to explore the data somewhat to understand how to clean it. In general, as you explore, you'll find things that need cleaning. As you clean, you can explore more. Therefore, at times it might make more sense to combine the cleaning and exploring code parts into a single R or Quarto file. Or split things in any other logical way.

As part of the exploratory analysis, you should produce plots or tables or other summary quantities for the most interesting/important quantities in your data. Depending on the total number of variables in your dataset, explore all or some of the others. Figures produced here might be histograms or density plots, correlation plots, etc. Tables might summarize your data.

Start by exploring one variable at a time. Then continue by creating plots or tables of the outcome(s) of interest and the predictor/exposure/input variables you are most interested in. If your dataset is small, you can do that for all variables.

Plots produced here can be scatterplots, boxplots, violinplots, etc. Tables can be simple 2x2 tables or larger ones.

# Setup

```{r}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
library(zoo)
```

Load the data.

```{r}
# importing data
mobility_data_2020 <- read.csv(here("data", "raw-data", "2020_US_Region_Mobility_Report.csv"))
mobility_data_2021 <- read.csv(here("data", "raw-data", "2021_US_Region_Mobility_Report.csv"))
mobility_data_2022 <- read.csv(here("data", "raw-data", "2022_US_Region_Mobility_Report.csv"))
combined_mobility_data <- bind_rows(mobility_data_2020, mobility_data_2021, mobility_data_2022)

time_series_covid <- read.csv(here("data", "raw-data", "time_series_covid19_confirmed_US.csv"))

chr_data_2024 <- read.csv(here("data", "raw-data", "chr_trends_csv_2024.csv"))

```

# Data exploration through graphs

Showing a bit of code to produce and save a summary table.

```{r}
# graph covid cases overtime

# Select only the date columns (starting from the 12th column)
date_columns_cases <- names(time_series_covid)[12:ncol(time_series_covid)]

# Ensure column names are correctly formatted as dates (removing 'X' from the date string)
date_fixed_cases <- as.Date(gsub("X", "", date_columns_cases), format = "%m.%d.%y")  # Fixed missing parenthesis

# Sum case counts across all locations for each date
total_cases <- colSums(time_series_covid[, date_columns_cases, drop = FALSE], na.rm = TRUE)

# Convert to a tidy data frame
total_cases_df <- data.frame(Date = date_fixed_cases, Cases = total_cases)

# Plot the total case counts over time with formatted y-axis labels without using 'scales'
case_count_graph <- ggplot(total_cases_df, aes(x = Date, y = Cases)) +
  geom_line(color = "blue") +
  scale_y_continuous(labels = function(x) format(x, big.mark = ",", scientific = FALSE)) +  # Format y-axis labels
  labs(title = "Total COVID-19 Cases Over Time in the U.S.",
       x = "Date",
       y = "Total Cases") +
  theme_minimal()

print(case_count_graph)
```

```{r}
# Calculate the daily new cases by subtracting previous day's total from the current day's total
new_cases <- c(0, diff(total_cases))  # diff() gives the difference between consecutive days; first day is set to 0

# Convert to a tidy data frame for new cases
new_cases_df <- data.frame(Date = date_fixed_cases, New_Cases = new_cases)

# Plot the new cases (incidence) over time with formatted y-axis labels
incidence_graph <- ggplot(new_cases_df, aes(x = Date, y = New_Cases)) +
  geom_line(color = "red") + 
  scale_y_continuous(labels = function(x) format(x, big.mark = ",", scientific = FALSE)) +  # Format y-axis labels
  labs(title = "Daily New COVID-19 Cases Over Time in the U.S.",
       x = "Date",
       y = "New Cases per Day") +
  theme_minimal()

print(incidence_graph)
```

```{r}
# Filter for U.S. data (adjust based on actual column name for region/country)
us_mobility_data <- combined_mobility_data %>%
  filter(country_region == "United States", sub_region_1 == "")

# Ensure the Date column is in Date format
us_mobility_data$Date <- as.Date(us_mobility_data$date, format = "%Y-%m-%d")  # Adjust if needed

# Calculate the 7-day moving average for each mobility type
us_mobility_data <- us_mobility_data %>%
  mutate(
    grocery_and_pharmacy_avg = rollmean(grocery_and_pharmacy_percent_change_from_baseline, 7, fill = NA, align = "right"),
    transit_stations_avg = rollmean(transit_stations_percent_change_from_baseline, 7, fill = NA, align = "right"),
    workplaces_avg = rollmean(workplaces_percent_change_from_baseline, 7, fill = NA, align = "right"),
    residential_avg = rollmean(residential_percent_change_from_baseline, 7, fill = NA, align = "right")
  )

# Plot the 7-day moving averages for each mobility type
mobility_graph <- ggplot(us_mobility_data, aes(x = Date)) +
  geom_line(aes(y = grocery_and_pharmacy_avg, color = "Grocery & Pharmacy")) +
  geom_line(aes(y = transit_stations_avg, color = "Transit Stations")) +
  geom_line(aes(y = workplaces_avg, color = "Workplaces")) +
  geom_line(aes(y = residential_avg, color = "Residential")) +
  labs(
    title = "7-Day Moving Average of Percent Change in Mobility in the U.S. Over Time",
    x = "Date",
    y = "7-Day Moving Average of Percent Change in Mobility",
    color = "Mobility Type"
  ) +
  theme_minimal()

print(mobility_graph)
```
```{r}

```

We are saving the results to the `results` folder. Depending on how many tables/figures you have, it might make sense to have separate folders for each. And/or you could have separate folders for exploratory tables/figures and for final tables/figures. Just choose a setup that makes sense for your project and works for you, and provide enough documentation that someone can understand what you are doing.

# Data exploration through figures

Histogram plots for the continuous outcomes.

Height first.

```{r}
p1 <- mydata %>% ggplot(aes(x=Height)) + geom_histogram() 
plot(p1)
figure_file = here("results", "figures","height-distribution.png")
ggsave(filename = figure_file, plot=p1) 
```

Now weights.

```{r}
p2 <- mydata %>% ggplot(aes(x=Weight)) + geom_histogram() 
plot(p2)
figure_file = here("results", "figures","weight-distribution.png")
ggsave(filename = figure_file, plot=p2) 
```

Now height as function of weight.

```{r}
p3 <- mydata %>% ggplot(aes(x=Height, y=Weight)) + geom_point() + geom_smooth(method='lm')
plot(p3)
figure_file = here("results", "figures", "height-weight.png")
ggsave(filename = figure_file, plot=p3) 
```

Once more height as function of weight, stratified by gender. Note that there is so little data, it's a bit silly. But we'll plot it anyway.

```{r}
p4 <- mydata %>% ggplot(aes(x=Height, y=Weight, color = Gender)) + geom_point() + geom_smooth(method='lm')
plot(p4)
figure_file = here("results", "figures", "height-weight-stratified.png")
ggsave(filename = figure_file, plot=p4) 
```

# Notes

For your own explorations, tables and figures can be "quick and dirty". As long as you can see what's going on, there is no need to polish them. That's in contrast to figures you'll produce for your final products (paper, report, presentation, website, etc.). Those should look as nice, polished and easy to understand as possible.