### 2025-02-21 code/eda-code/eda.qmd
Project Part 2
Might be a bit bare, however, I took caution to fleshing out too much if my topic needs to be adjusted. Right now, I have considered adding in the CDC Vaccination Hesitancy survey (possibly to map out and compare with mobility dynamics) or the County Health Ranking (possibly to explore how communities of differing SES dealt with cases in terms of mobility).

---
title: "Exploratory Data Analysis"
date: "2025-02-21"
output: html_document
---

This Quarto file loads the cleaned data and does some exploring.
# Setup

```{r}
#load needed packages. make sure they are installed
library(here) 
library(dplyr)
library(skimr)
library(ggplot2)
library(zoo)
library(tidyr)
library(tigris)
library(sf)
```

Load the data from processingfile-v1 file.

```{r}

# processed data
processeddata <- readRDS(here("data", "processed-data", "processeddata.rds"))

```

Creation of a heatmap showing total COVID 19 cases on the last observed date, 10/15/2022. I also created a heat map showing mobility changes, with hotter indicating more mobility to retail and recreational spots. Interestingly, the two maps seem to correlate even though I used total COVID-19 cases rather than new ones.
```{r}
# Load the state-level shapefile 
us_map_states <- tigris::states(cb = TRUE) %>%
  st_as_sf()  # Convert the shapefile to a Simple Feature object

# Clean the state names in the shapefile
us_map_clean <- us_map_states %>%
  mutate(NAME = tolower(NAME))  # Convert state abbreviation to lowercase

# Filter your dataset to get data for the last recorded day and group by state
last_day <- max(processeddata$date)
data_last_day <- processeddata %>%
  filter(date == last_day) %>%
  group_by(state) %>%
  summarise(total_cases = sum(total_cases, na.rm = TRUE),  # Sum of total cases
            retail_and_recreation_percent_change_from_baseline = 
              mean(retail_and_recreation_percent_change_from_baseline, na.rm = TRUE)) %>%  # Average mobiltiy changes
  mutate(state = tolower(state))  # Convert state names to lowercase

# Join the shapefile with the COVID-19 data
us_map_data <- us_map_clean %>%
  left_join(data_last_day, by = c("NAME" = "state"))

# plot total COVID-19 cases to USA map on 10/15/2022
cases_map <- ggplot(us_map_data) +
  geom_sf(aes(fill = total_cases), color = "white") +
  scale_fill_viridis_c(option = "plasma", trans = "log", labels = NULL) + 
  labs(title = paste("Heatmap of Total COVID-19 Cases in the U.S. on", last_day),
       fill = "Hotter = More Cases") +
  theme_minimal() +
  coord_sf(xlim = c(-125, -65), ylim = c(25, 50))

print(cases_map)

figure_file = here("results","figures","casemap.png")
ggsave(filename = figure_file, plot=cases_map)

# Create the heatmap for retail and recreation mobility changes
mobility_map <- ggplot(us_map_data) +
  geom_sf(aes(fill = retail_and_recreation_percent_change_from_baseline), color = "white") +  
  scale_fill_viridis_c(option = "inferno", trans = "identity") +
  labs(title = paste("Heatmap of Retail and Recreation Mobility Changes in the U.S. on", last_day),
       fill = "Mobility Change") +
  theme_minimal() +
  theme(
        legend.spacing.y = unit(1, "cm"),  
        legend.box.spacing = unit(1, "cm"),  
        legend.key.size = unit(1, "cm")) +  
  coord_sf(xlim = c(-125, -65), ylim = c(25, 50))

print(mobility_graph)

# save graph
figure_file = here("results","figures","mobilitymap.png")
ggsave(filename = figure_file, plot=mobility_map)
```

This graph aims to plot total cases overtime for every location on a given day. This required the creation of a dataframe to store the sum of all total cases.
```{r}
# Create new data frame to have cases per day instead of state/county
total_cases_df <- processeddata %>%
  group_by(date) %>%
  summarise(Cases = sum(total_cases, na.rm = TRUE)) %>%
  ungroup()

# Plot the total case counts over time
case_count_graph <- ggplot(total_cases_df, aes(x = date, y = Cases)) +
  geom_line(color = "blue") +
  scale_y_continuous(labels = function(x) format(x, big.mark = ",", scientific = FALSE)) +  
  labs(title = "Total COVID-19 Cases Over Time in the U.S.",
       x = "Date",
       y = "Total Cases") +
  theme_minimal()

print(case_count_graph)

# save graph
figure_file = here("results","figures","casecountplot.png")
ggsave(filename = figure_file, plot=case_count_graph)
```
This graph aims to plot new cases overtime for every location on a given day. This required the creation of a dataframe to store the sum of all new cases.
```{r}
# data frame to sum all cases on a given day for every location
new_cases_df <- processeddata %>%
  group_by(date) %>%
  summarise(Cases = sum(new_cases, na.rm = TRUE)) %>%
  ungroup()

# Plot the new cases over time
incidence_graph <- ggplot(new_cases_df, aes(x = date, y = Cases)) +
  geom_line(color = "red") + 
  scale_y_continuous(labels = function(x) format(x, big.mark = ",", scientific = FALSE)) +
  labs(title = "Daily New COVID-19 Cases Over Time in the U.S.",
       x = "Date",
       y = "New Cases per Day") +
  theme_minimal()

print(incidence_graph)

# save graph
figure_file = here("results","figures","incidenceplot.png")
ggsave(filename = figure_file, plot=incidence_graph)

```
I aim to plot out the changes in mobility for each category as time progresses. Originally, it plotted out everyday, however, the line was far too jagged so I applied a weekly rolling average instead. I also had to create a dataframe to average values for a given day for every location.
```{r}

# Create dataframe of average mobility for everywhere in the US per date
us_average_mobility <- processeddata %>%
  group_by(date) %>%
  summarise(
    total_mobility_changes = mean(average_mobility, na.rm = TRUE),
    transit_mean = mean(transit_stations_percent_change_from_baseline, na.rm = TRUE),
    parks_mean = mean(parks_percent_change_from_baseline, na.rm = TRUE),
    retail_mean = mean(retail_and_recreation_percent_change_from_baseline, na.rm = TRUE),
    grocery_mean = mean(grocery_and_pharmacy_percent_change_from_baseline, na.rm = TRUE),
    workplaces_mean = mean(workplaces_percent_change_from_baseline, na.rm = TRUE),
    residential_mean = mean(residential_percent_change_from_baseline, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  arrange(date) %>%  # Ensure dates are sorted
  mutate(
    transit_mean_7day = rollmean(transit_mean, k = 7, fill = NA, align = "right"),
    parks_mean_7day = rollmean(parks_mean, k = 7, fill = NA, align = "right"),
    retail_mean_7day = rollmean(retail_mean, k = 7, fill = NA, align = "right"),
    grocery_mean_7day = rollmean(grocery_mean, k = 7, fill = NA, align = "right"),
    workplaces_mean_7day = rollmean(workplaces_mean, k = 7, fill = NA, align = "right"),
    residential_mean_7day = rollmean(residential_mean, k = 7, fill = NA, align = "right")
  )

# Plot the 7-day averages for each mobility type
mobility_graph <- ggplot(us_average_mobility, aes(x = date)) +
  geom_line(aes(y = grocery_mean_7day, color = "Grocery & Pharmacy")) +
  geom_line(aes(y = transit_mean_7day, color = "Transit Stations")) +
  geom_line(aes(y = workplaces_mean_7day, color = "Workplaces")) +
  geom_line(aes(y = residential_mean_7day, color = "Residential")) +
  geom_line(aes(y = parks_mean_7day, color = "Recreational")) +
  labs(
    title = "7-Day Moving Average of Percent Change in Mobility in the U.S. Over Time",
    x = "Date",
    y = "7-Day Average of Percent Change in Mobility",
    color = "Mobility Type"
  ) +
  theme_minimal()

print(mobility_graph)

# save graph
figure_file = here("results","figures","mobilitygraph.png")
ggsave(filename = figure_file, plot=mobility_graph)
```