### 2025-02-21 code/processing-code/processingfile-v1.qmd

---
title: "Cleaning Script"
author: "Vincent Nguyen"
date: "2025-02-21"
output: html_document
---

# Setup

Load needed packages. make sure they are installed.

```{r}
# load needed packages
library(here) 
library(dplyr)
library(skimr)
library(ggplot2)
library(zoo)
library(tidyr)
```

# Data loading

```{r}
# importing data

# google data is in three parts
mobility_data_2020 <- read.csv(here("data", "raw-data", "2020_US_Region_Mobility_Report.csv"))
mobility_data_2021 <- read.csv(here("data", "raw-data", "2021_US_Region_Mobility_Report.csv"))
mobility_data_2022 <- read.csv(here("data", "raw-data", "2022_US_Region_Mobility_Report.csv"))
combined_mobility_data <- bind_rows(mobility_data_2020, mobility_data_2021, mobility_data_2022)

# jhk covid time series cases
time_series_covid <- read.csv(here("data", "raw-data", "time_series_covid19_confirmed_US.csv"))

# county health ranking
chr_data_2024 <- read.csv(here("data", "raw-data", "chr_trends_csv_2024.csv"))

```

After visually inspecting the mobility data, I noticed that many mobility changes were missing. The following code seeks to quantify this.
```{r}

# After visually inspecting the data, I noticed that a lot of mobility changes were missing.
sapply(combined_mobility_data[c("transit_stations_percent_change_from_baseline", 
                         "parks_percent_change_from_baseline", 
                         "retail_and_recreation_percent_change_from_baseline", 
                         "grocery_and_pharmacy_percent_change_from_baseline",
                         "workplaces_percent_change_from_baseline", 
                         "residential_percent_change_from_baseline")], function(x) sum(is.na(x)))

```
# Cleaning and Wrangling

By inspecting the data, some problems become apparent. AS mentioned above, the Mobility Report is missing a tremendous amount of values. The Time Series data and Mobility Report are structured in entirely different ways. Joining the two will be difficult without mutating one. I chose to change the Time Series data.

Next, we also can see that the naming scheme of the columns differ so we need to standardize that as well.

Lastly, both datasets do not have the same range in dates

```{r}

# Pivot the time series data and change names around
cases_long <- time_series_covid %>%
  pivot_longer(
    cols = starts_with("X"),  
    names_to = "date",
    values_to = "total_cases"
  ) %>%
  mutate(date = gsub("^X", "", date),  # Remove leading x for dates to join data sets in a bit
         date = gsub("\\.", "/", date),  # Replace dots with slashes
         date = as.Date(date, format = "%m/%d/%y"))  # Use "%y" instead of "%Y"

# Standardize county & state names (convert to lowercase to avoid mismatches)
cases_long <- cases_long %>%
  rename(county = Admin2, state = Province_State) %>%
  mutate(county = tolower(county), state = tolower(state))

mobility_df <- combined_mobility_data %>%
  rename(county = sub_region_2, state = sub_region_1) %>%
  mutate(county = tolower(county), 
         state = tolower(state),
         county = gsub(" county$", "", county),  # Remove "county" from county name
         date = as.Date(date, format = "%Y-%m-%d"))  # Correct format for "YYYY-MM-DD"

# Filter cases_long and mobility_df based on the date range
cases_long_filtered <- cases_long %>%
  filter(date >= "2020-02-15" & date <= "2022-10-15")

mobility_df_filtered <- mobility_df %>%
  filter(date >= "2020-02-15" & date <= "2022-10-15")

# Merge the data sets together
merged_df <- cases_long_filtered %>%
  inner_join(mobility_df_filtered, by = c("county", "state", "date"))

# Finally, remove any rows missing mobility changes to ensure consistency. We also have a lot of data values so I deemed it permissible.
merged_df_clean <- merged_df %>%
  drop_na(c(transit_stations_percent_change_from_baseline, parks_percent_change_from_baseline, retail_and_recreation_percent_change_from_baseline, grocery_and_pharmacy_percent_change_from_baseline,workplaces_percent_change_from_baseline, residential_percent_change_from_baseline))

# Additionally, add a column about new cases for further analysis down the line.
merged_df_clean <- merged_df_clean %>%
  arrange(state, county, date) %>%  # Sort (Need to make sure calculations occur between the correct states/counties and in order)
  group_by(state, county) %>%  # Group by state and county to calculate new cases within each county
  mutate(new_cases = c(NA, diff(total_cases))) %>%  # Calculate new cases
  ungroup()  # Ungroup after the calculation
```

All done, data is clean now.

Let's assign at the end to some final variable, this makes it easier to add further cleaning steps above.

```{r}
processeddata <- merged_df_clean
```

# Save data

```{r}
save_data_location <- here::here("data","processed-data","processeddata.rds")
saveRDS(processeddata, file = save_data_location)
```