### 2025-04-04 code/processing-code/processingfile.qmd

---
title: "Cleaning Script"
author: "Vincent Nguyen"
date: "2025-02-21"
output: html_document
---

# Setup

Load needed packages. make sure they are installed.

```{r}
# load needed packages
library(here) 
library(dplyr)
library(skimr)
library(ggplot2)
library(zoo)
library(tidyr)
library(lubridate)
library(readxl)
library(stringr)
```

# Data loading

```{r}
# google data is in three parts
mobility_data_2020 <- read.csv(here("data", "raw-data", "2020_US_Region_Mobility_Report.csv")) %>%
  filter(sub_region_1 == "Georgia")
mobility_data_2021 <- read.csv(here("data", "raw-data", "2021_US_Region_Mobility_Report.csv"))%>%
  filter(sub_region_1 == "Georgia")
mobility_data_2022 <- read.csv(here("data", "raw-data", "2022_US_Region_Mobility_Report.csv"))%>%
  filter(sub_region_1 == "Georgia")

combined_mobility_data <- bind_rows(mobility_data_2020, mobility_data_2021, mobility_data_2022)

# jhk covid time series cases
time_series_covid <- read.csv(here("data", "raw-data", "time_series_covid19_confirmed_US.csv")) %>%
  filter(Province_State == "Georgia")

# county health ranking
#hr_data_2024 <- read.csv(here("data", "raw-data", "chr_trends_csv_2024.csv")) %>%
#  filter(state == "GA")

# county population information
ch_pop_data <- read_excel(here::here("data", "raw-data", "co-est2023-pop.xlsx"), skip = 3)

# Load in CHR Data from 2020 to 2022, remove the first column, remove first two rows, add a year column, and also rename the columns

chr_2020 <- read_excel(here::here("data", "raw-data", "2020 County Health Rankings Georgia Data - v1_0.xlsx"), sheet = 2, skip = 2) %>% select(-1) %>% mutate(year = 2020)

chr_2020 <- chr_2020 %>%
  rename(
    state = `Georgia`,
    county = `...3`,
    health_outcome_zscore = `...4`,
    health_outcome_rank = `...5`,
    health_factor_zscore = `...6`,
    health_factor_rank = `...7`
  )

chr_2021 <- read_excel(here::here("data", "raw-data", "2021 County Health Rankings Georgia Data - v1.xlsx"), sheet = 2, skip = 2)%>% select(-1) %>% mutate(year = 2021)

chr_2021 <- chr_2021 %>%
  rename(
    state = `Georgia`,
    county = `...3`,
    health_outcome_zscore = `...4`,
    health_outcome_rank = `...5`,
    health_factor_zscore = `...6`,
    health_factor_rank = `...7`
  )

chr_2022 <- read_excel(here::here("data", "raw-data", "2022 County Health Rankings Georgia Data - v2.xlsx"), sheet = 2, skip = 2)%>% select(-1) %>% mutate(year = 2022)

chr_2022 <- chr_2022 %>%
  rename(
    state = `Georgia`,
    county = `...3`,
    health_outcome_zscore = `...4`,
    health_outcome_rank = `...5`,
    health_factor_zscore = `...6`,
    health_factor_rank = `...7`
  )

# Combine all CHR data sets into one
combined_chr <- bind_rows(chr_2020, chr_2021, chr_2022)
```

```{r}
# Remove the second column
ch_pop_data <- ch_pop_data[, -2]

ch_pop_data <- ch_pop_data %>%
  pivot_longer(
    cols = c(`2020`,`2021`, `2022`, `2023`),    
    names_to = "year",        
    values_to = "population_count"              
  )

# Fix the first column
ch_pop_data <- ch_pop_data %>%
  mutate(
    county = str_remove(`...1`, "^\\."),            
    county = str_remove(county, "County"),     
    county = str_to_lower(county),                  
    state = str_extract(county, "(?<=,).*"),        
    county = str_remove(county, ",.*")              
  ) %>%
  mutate(
    state = str_to_lower(str_trim(state))           
  )

# Change year column to be double for joining later
ch_pop_data$year <- as.double(ch_pop_data$year)

# Clean names to prepare for joining
ch_pop_data <- ch_pop_data %>%
  mutate(
    county = str_trim(str_to_lower(county)),  
    state = str_trim(str_to_lower(state))     
  )

# Only include Georgia counties
ch_pop_data <- ch_pop_data %>%
  filter(state == "georgia")

```

# Cleaning and Wrangling

By inspecting the data, some problems become apparent. AS mentioned above, the Mobility Report is missing a tremendous amount of values. The Time Series data and Mobility Report are structured in entirely different ways. Joining the two will be difficult without mutating one. I chose to change the Time Series data.

Next, we also can see that the naming scheme of the columns differ so we need to standardize that as well.

Lastly, both datasets do not have the same range in dates

```{r}

# Pivot the time series data and change names around
cases_long <- time_series_covid %>%
  pivot_longer(
    cols = starts_with("X"),  
    names_to = "date",
    values_to = "total_cases"
  ) %>%
  mutate(date = gsub("^X", "", date),  # Remove leading x for dates to join data sets in a bit
         date = gsub("\\.", "/", date),  # Replace dots with slashes
         date = as.Date(date, format = "%m/%d/%y"))  # Use "%y" instead of "%Y"

# Standardize county & state names (convert to lowercase to avoid mismatches)
cases_long <- cases_long %>%
  rename(county = Admin2, state = Province_State) %>%
  mutate(county = tolower(county), state = tolower(state))

mobility_df <- combined_mobility_data %>%
  rename(county = sub_region_2, state = sub_region_1) %>%
  mutate(county = tolower(county), 
         state = tolower(state),
         county = gsub(" county$", "", county),  # Remove "county" from county name
         date = as.Date(date, format = "%Y-%m-%d"))  # Correct format for "YYYY-MM-DD"

# Filter cases_long and mobility_df based on the date range
cases_long_filtered <- cases_long %>%
  filter(date >= "2020-02-15" & date <= "2022-10-15")

mobility_df_filtered <- mobility_df %>%
  filter(date >= "2020-02-15" & date <= "2022-10-15")

# Merge the data sets together
merged_df <- cases_long_filtered %>%
  inner_join(mobility_df_filtered, by = c("county", "state", "date"))

# Finally, remove any rows missing mobility changes to ensure consistency. We also have a lot of data values so I deemed it permissible.
#merged_df_clean <- merged_df %>%
#  drop_na(c(transit_stations_percent_change_from_baseline, parks_percent_change_from_baseline, retail_and_recreation_percent_change_from_baseline, grocery_and_pharmacy_percent_change_from_baseline,workplaces_percent_change_from_baseline, residential_percent_change_from_baseline))
merged_df_clean <- merged_df
# Additionally, add a column about new cases for further analysis down the line.
merged_df_clean <- merged_df_clean %>%
  arrange(state, county, date) %>%  
  group_by(state, county) %>%  
  mutate(new_cases = c(NA, diff(total_cases))) %>%  
  ungroup()

# Create column for mean average mobiltiy change for that given day and location. I omitted mobility changes to residencies because that may be more of an indication of who is at home.
merged_df_clean <- merged_df_clean %>%
  arrange(state, county, date) %>%  # Sort to ensure correct calculations
  group_by(state, county) %>%  # Group by state and county
  mutate(average_mobility = rowMeans(
    cbind(transit_stations_percent_change_from_baseline, 
          parks_percent_change_from_baseline, 
          retail_and_recreation_percent_change_from_baseline, 
          grocery_and_pharmacy_percent_change_from_baseline, 
          workplaces_percent_change_from_baseline), 
    na.rm = TRUE)) %>%
  ungroup()

merged_df_clean <- merged_df_clean %>%
  mutate(year = year(date)) %>%
  left_join(
    combined_chr %>%
      mutate(across(c(state, county), str_to_lower)),  # Convert state and county to lowercase
    by = c("state", "county", "year")
  )

merged_df_clean <- merged_df_clean %>%
  left_join(
    ch_pop_data %>% 
      select(year, population_count, county, state),
    by = c("state" = "state", "county" = "county", "year" = "year")
  )

# Create column for incidence rate per 10,000 (10,000 because numbers tend to be small when per 100k)
merged_df_clean <- merged_df_clean %>%
  mutate(incidence_rate = (new_cases / population_count) * 10000)

# Clean up and Removal of unnecessary columns
merged_df_clean <- merged_df_clean %>% select(-UID, -iso2, -iso3, -code3, -country_region_code, -country_region, -metro_area, -iso_3166_2_code, -census_fips_code, -place_id)
```

This code chunk entails the creation of a population density column in the data.

```{r}
library(tigris)
library(sf)

# Load county shapefile for GA
ga_counties <- counties(state = "GA", cb = TRUE, class = "sf")

# Make county column lowercase for merging
ga_counties <- ga_counties %>%
  mutate(NAME = tolower(NAME))

# Merge with processed population data (assumed already cleaned)
ga_counties <- ga_counties %>%
  left_join(ch_pop_data, by = c("NAME" = "county"))

# Calculate county area in square kilometers
ga_counties <- ga_counties %>%
  mutate(area_km2 = as.numeric(st_area(geometry)) / 1e6)  # Convert m² to km²

# Compute population density (people per km²) and remove unnecessary data
ga_counties <- ga_counties %>%
  mutate(pop_density = population_count / area_km2) %>%
  select(pop_density, year, NAME)

# Add in population density column
merged_df_clean <- merged_df_clean %>%
  left_join(ga_counties, by = c("county" = "NAME", "year" = "year"))
```
```{r}
# Clean negative and NA values
merged_df_clean$new_cases[merged_df_clean$new_cases < 0] <- 0
merged_df_clean$new_cases[is.na(merged_df_clean$new_cases)] <- 0
merged_df_clean$incidence_rate[is.na(merged_df_clean$incidence_rate)] <- 0

# Remove any remaining NA values
merged_df_clean <- na.omit(merged_df_clean)

# Filter for pandemic period
last_case_date <- as.Date("2022-01-01")
first_case_date <- as.Date("2020-03-02")

merged_df_clean <- merged_df_clean %>%
  filter(date >= first_case_date) %>%
  filter(date <= last_case_date) %>%
  group_by(county)

# Convert to tibble and remove geometry
merged_df_clean <- merged_df_clean %>%
  as_tibble() %>%
  select(-geometry)

# Convert new_cases to integer
merged_df_clean <- merged_df_clean %>% 
  mutate(new_cases = as.integer(new_cases))
```

All done, data is clean now.

Let's assign at the end to some final variable, this makes it easier to add further cleaning steps above.

```{r}
processeddata <- merged_df_clean
```

# Save data

```{r}
save_data_location <- here::here("data","processed-data","processeddata.rds")
saveRDS(processeddata, file = save_data_location)
```