### 2025-04-04 code/analysis-code/statisitical-analysis.qmd
Fixes

---
title: "Analysis Script"
author: "Vincent Nguyen"
date: "2025-02-21"
output: html_document
---

# Setup

Load needed packages. make sure they are installed.

```{r}
library(ggplot2) 
library(broom) 
library(here) 
library(glmnet)
library(MASS)
library(tidymodels)
library(dplyr)
library(rsample)
library(poissonreg)
library(parsnip)
library(future)
```

```{r}
#path to data
#note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","processeddata.rds")

#load data. 
data <- readRDS(data_location)

# NOTES TO SELF
# - change to QMD file (personally prefer)
# - work on other analysis (lag-time & quarantine fatigue? more research)
# - work on inclusion of county health ranking data or possibly not include at all? lots of mismatch counties between datasets
class(data$new_cases)
```

```{r}
# Intial Recipe setting
recipe <- recipe(new_cases ~ pop_density + transit_stations_percent_change_from_baseline + 
                     parks_percent_change_from_baseline + 
                     retail_and_recreation_percent_change_from_baseline + 
                     grocery_and_pharmacy_percent_change_from_baseline +
                     workplaces_percent_change_from_baseline + 
                     residential_percent_change_from_baseline +
                     health_factor_rank, data = data)

```

First, I created a linear regression model as the baseline standard. This model performs very poorly as it has a low R-squared, indicating low predictive power. The model does mark all variables of the model to be significant.Additionally, the AIC is very extremely high.

```{r}
# Create linear regression model
lm_spec <- linear_reg() %>%
  set_engine("lm")

lm_wf <- workflow() %>%
  add_recipe(recipe)

lm_fit <- lm_wf %>%
  add_model(lm_spec) %>%
  fit(data = data)

lm_fit %>%
  extract_fit_parsnip() %>%
  tidy()

lm_fit %>%
  extract_fit_parsnip() %>%
  broom::glance()

lm_preds<- predict(lm_fit, new_data = data)

lm_metrics <- tibble(truth = data$new_cases, predicted = lm_preds$.pred) %>%
  metrics(truth = truth, estimate = predicted)
lm_metrics
```

I utilized Lasso here to hopefully bring down the RMSE and bring up the rsquared. It seems Lasso has no effect on model performance as the data lacks multicolinearity and does not have irrelevant predictors. 
```{r}
# Set method for LASSO
lasso_spec <- linear_reg(penalty = 0.1, mixture = 1) %>%
  set_engine("glmnet")

# Insert recipe
lasso_wf <- workflow() %>%
  add_recipe(recipe)

lasso_fit <- lasso_wf %>%
  add_model(lasso_spec) %>%
  fit(data = data)

lasso_fit %>%
  extract_fit_parsnip() %>%
  tidy()

lasso_fit %>%
  extract_fit_parsnip() %>%
  broom::glance()

lasso_preds<- predict(lasso_fit, new_data = data)

lasso_metrics <- tibble(truth = data$new_cases, predicted = lasso_preds$.pred) %>%
  metrics(truth = truth, estimate = predicted)
lasso_metrics
```

After the last two models performed poorly, the use of random forest seemed rational as this model can capture more complex relationships. 
```{r}

rngseed = 1234

forest_spec <- rand_forest(mode = "regression") %>%
  set_engine("ranger", seed = rngseed)

# Insert recipe
forest_wf <- workflow() %>%
  add_recipe(recipe)

# Create model with all predictors and random forest
forest_fit <- forest_wf %>%
 add_model(forest_spec) %>%
  fit(data = data)

forest_preds <- predict(forest_fit, new_data = data)
forest_preds <- tibble(truth = data$new_cases, predicted = forest_preds$.pred)
forest_rmse <- forest_preds %>%
  metrics(truth = truth, estimate = predicted)
print(forest_rmse)
```

```{r}
# Tuning forest
forest_spec <- rand_forest(
  mode = "regression", 
  trees = 300, 
  mtry = tune(), 
  min_n = tune()
) %>%
  set_engine("ranger")

# Create grid with set parameters
forest_grid <- grid_regular(
  mtry(range = c(1, 7)),  
  min_n(range = c(1, 21)),  
  levels = 7  
)

# Set seed
set.seed(1234)

# 5 fold crossvalidation repeated 5 times
# Note: Limitation of computing power so doing rolling window would not work
cv_folds <- vfold_cv(data, v = 5, repeats = 5)
plan(multisession, workers = 4)

# Tune the forest
# This step took 20-ish minutes
forest_tune_res <- tune_grid(
  forest_wf %>% add_model(forest_spec),
  resamples = cv_folds,  
  grid = forest_grid,
  metrics = metric_set(rmse, rsq),
  control = control_grid(parallel_over = "everything")
)

# Plot results
autoplot(forest_tune_res)

forest_tune_res %>%
  collect_metrics()

best_params <- select_best(forest_tune_res, metric = "rmse")
best_params

best_params$trees <- 500
```

```{r}
forest_spec <- rand_forest(
  mode = "regression",
  mtry = best_params$mtry,
  trees = best_params$trees,
  min_n = best_params$min_n
) %>%
  set_engine("ranger", seed = 1234)

forest_fit_final <- forest_wf %>%
  add_model(forest_spec) %>%
  fit(data = data)

# Get final predictions
forest_preds_final <- predict(forest_fit_final, new_data = data)
forest_preds_final <- tibble(truth = data$new_cases, predicted = forest_preds_final$.pred)

# Calculate performance metrics
forest_rmse <- forest_preds_final %>%
  metrics(truth = truth, estimate = predicted)

print(forest_rmse)
```