### 2025-04-18 code/eda-code/eda.qmd

---
title: "Exploratory Data Analysis"
date: "2025-02-21"
output: html_document
---

This Quarto file loads the cleaned data and does some exploring. \# Setup

```{r}
#load needed packages. make sure they are installed
library(here) 
library(dplyr)
library(ggplot2)
library(zoo)
library(patchwork)
library(e1071)
library(scales)
library(RColorBrewer)
library(corrplot)
library(reshape2)

```

Load the data from processingfile-v1 file.

```{r}

# processed data
data <- readRDS(here("data", "processed-data", "processeddata.rds"))

case_data <- readRDS(here("data", "processed-data", "processedcasedata.rds"))

data<- data %>%
   mutate(incidence_rate = if_else(incidence_rate < 0, 0, incidence_rate))
```

DataVisualization

Since new case incidence is the variable of interest, it was visually inspected to understand the distribution.

From the graphs, kurtosis test, and skewness test, it is apparent that daily new case incidence is not normally distribued and heavily skewed towards 0 and other lower values. The second graph, new cases over 500, was graphed to show how certain days do reach high case incidence.
```{r}

density_cases <- data %>%
  ggplot( aes(x=new_cases)) +
    geom_density(fill="#69b3a2", color="#e9ecef", alpha=0.8) +
  labs(title = "Distribution of Daily New Case Incidence", 
       x = "Daily New Cases",
       y = "Density") +
  theme_bw()

print(density_cases)

figure_file = here("results","figures","density_cases.png")
ggsave(filename = figure_file, plot=density_cases)

density_cases_500 <- data %>%
  filter(new_cases > 500) %>%
  ggplot( aes(x=new_cases)) +
    geom_density(fill="#69b3a2", color="#e9ecef", alpha=0.8) +
  labs(title = "Distribution of Daily New Case Incidence (Over 500)", 
       x = "Daily New Cases",
       y = "Density") +
  theme_bw()

print(density_cases_500)

figure_file = here("results","figures","density_cases_500.png")
ggsave(filename = figure_file, plot=density_cases_500)

print(kurtosis(data$new_cases))
print(skewness(data$new_cases))
```
A histogram of population density was created. I hypothesize that population density plays a large role in case incidence. Population density is also skewed, however, there are some notable counties that do have large populationd densities.
```{r}
data_summary <- data %>%
  group_by(county) %>%
  summarize(pop_density = mean(pop_density, na.rm = TRUE))

histo_pop_density <- ggplot(data_summary, aes(x = pop_density)) + 
  geom_histogram(bins = 30, fill = "#69b3a2", color = "black", alpha = 0.7) +
  labs(title = "Histogram of Population Density (2-Year Average per County)", 
       x = "Population Density (Calculated as Population Count / km squared)", y = "Frequency") + theme_bw()

print(histo_pop_density)

figure_file = here("results","figures","histo_pop_density.png")
ggsave(filename = figure_file, plot=histo_pop_density)
```

A 7-day rolling average was applied to the mobility data as the original iteration was too jagged. This code entails the creation of a data frame for 7 day averages.

```{r}

# Create dataframe of average mobility per county per date
ga_average_mobility <- data %>%
  group_by(date) %>%
  summarise(
    retail_mean = mean(retail_and_recreation_percent_change_from_baseline, na.rm = TRUE),
    grocery_mean = mean(grocery_and_pharmacy_percent_change_from_baseline, na.rm = TRUE),
    workplaces_mean = mean(workplaces_percent_change_from_baseline, na.rm = TRUE),
    residential_mean = mean(residential_percent_change_from_baseline, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  arrange(date) %>%  # Ensure dates are sorted
  mutate(
    retail_mean_7day = rollmean(retail_mean, k = 7, fill = NA, align = "right"),
    grocery_mean_7day = rollmean(grocery_mean, k = 7, fill = NA, align = "right"),
    workplaces_mean_7day = rollmean(workplaces_mean, k = 7, fill = NA, align = "right"),
    residential_mean_7day = rollmean(residential_mean, k = 7, fill = NA, align = "right")
  )

# Plot the 7-day averages for each mobility type
mobility_graph <- ggplot(ga_average_mobility, aes(x = date)) +
  geom_line(aes(y = grocery_mean_7day, color = "Grocery & Pharmacy")) +
  geom_line(aes(y = workplaces_mean_7day, color = "Workplaces")) +
  geom_line(aes(y = residential_mean_7day, color = "Residential")) +
  geom_line(aes(y = retail_mean_7day, color = "Retail")) +
  labs(
    title = "7-Day Moving Average of Percent Change in Mobility in Georgia Over Time",
    x = "Date",
    y = "7-Day Average of Percent Change in Mobility",
    color = "Mobility Type"
  ) +
  theme_bw()

print(mobility_graph)

# save graph
figure_file = here("results","figures","mobilitygraph.png")
ggsave(filename = figure_file, plot=mobility_graph)
```

For better legiiblity, all 4 mobiltiy types were graphed separately.

```{r}
# Create one plot per mobility type
p1 <- ggplot(ga_average_mobility, aes(x = date, y = grocery_mean_7day)) +
  geom_line(color = "steelblue") +
  labs(title = "Grocery & Pharmacy", x = "Date", y = "7-Day Avg % Change") +
  theme_bw() +
  coord_cartesian(ylim = c(-50, 20))

p2 <- ggplot(ga_average_mobility, aes(x = date, y = workplaces_mean_7day)) +
  geom_line(color = "darkred") +
  labs(title = "Workplaces", x = "Date", y = "7-Day Avg % Change") +
  theme_bw() +
  coord_cartesian(ylim = c(-50, 20))

p3 <- ggplot(ga_average_mobility, aes(x = date, y = residential_mean_7day)) +
  geom_line(color = "forestgreen") +
  labs(title = "Residential", x = "Date", y = "7-Day Avg % Change") +
  theme_bw() +
  coord_cartesian(ylim = c(-50, 20))

p4 <- ggplot(ga_average_mobility, aes(x = date, y = retail_mean_7day)) +
  geom_line(color = "purple") +
  labs(title = "Retail", x = "Date", y = "7-Day Avg % Change") +
  theme_bw() +
  coord_cartesian(ylim = c(-50, 20))

# Arrange in a 2x2 grid
mobility_grid <- (p1 | p2) / (p3 | p4)

# Show it
print(mobility_grid)

# Save it
figure_file = here("results", "figures", "mobility_grid.png")
ggsave(filename = figure_file, plot = mobility_grid)

```

After visually inspecting the new cases over time graph, heteroscedastity and non-stationairty are of concern. Large outbreaks can skew these metrics and even when tested, could be misleading. Additionally, assumptions of homoscedastity are required for some tests of stationairty and some assumptions of normality are required for some tests of homoscdastity. Because of these concerns, I chose to focus on models like XGBoost and Random Forest which do not require these assumptions.

```{r}
library(scales)

first_date <- min(data$date)
last_date <- max(data$date)

cases_graph <- case_data %>%
  group_by(date) %>%
  summarise(total_cases = sum(total_cases, na.rm = TRUE)) %>%
  arrange(date) %>%
  ggplot(aes(x = date, y = total_cases)) +
    geom_area(fill = "#69b3a2", alpha = 0.5) +
    geom_line(color = "#69b3a2") +
    ylab("Total Cases (in GA)") +
    scale_y_continuous(labels = label_number()) +
   scale_x_date(
      breaks = seq(from = first_date, to = last_date, by = "6 months"),  # Breaks every 3 months
      labels = date_format("%b %Y")  # Format labels as month year (e.g., Jan 2020)
    ) +
    theme_bw() +
  labs(title = "Total Cases (in GA) Over Time",
       x = "Date (2020 - 2022)",
       y = "Total Cases")

print(cases_graph)

figure_file = here("results","figures","cases_graph.png")
ggsave(filename = figure_file, plot=cases_graph)

incidence_graph <- data %>%
  group_by(date) %>%
  summarise(new_cases = sum(new_cases, na.rm = TRUE)) %>%
  arrange(date) %>%
  ggplot(aes(x = date, y = new_cases)) +
    geom_area(fill = "#69b3a2", alpha = 0.5) +
    geom_line(color = "#69b3a2") +
    ylab("Daily New Cases") +
     scale_x_date(
      date_breaks = "3 months",  # Automatically set the breaks to every 3 months
      date_labels = "%b %Y"      # Format labels as month-year (e.g., Jan 2020)
    ) +
    theme_bw() +
  labs(title = "Daily New Cases (in GA) Over Time",
       x = "Date (2020 - 2022)",
       y = "Total Cases")

print(incidence_graph)

figure_file = here("results","figures","incidence_graph.png")
ggsave(filename = figure_file, plot=incidence_graph)
```

This code chunk plots a scatter plot of a county's population count against its final total cases. 
```{r}

year_end_cases <- data %>%
  mutate(year = year(date)) %>%
  group_by(county, year) %>%
  slice_max(order_by = date, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(county, year, total_cases)

pop_density_info <- data %>%
  mutate(year = year(date)) %>%
  select(county, year, pop_density) %>%
  distinct()

plot_df <- left_join(year_end_cases, pop_density_info, by = c("county", "year"))

year_end_cases_graph <- ggplot(plot_df, aes(x = pop_density, y = total_cases)) +
  geom_point(color = "#69b3a2", alpha = 0.6) +
  facet_wrap(~year) +
  labs(
    x = "Population Density",
    y = "Total COVID Cases (End of Year)",
    title = "Year-End COVID Cases vs. Population Density by County"
  ) +
  theme_bw()

print(year_end_cases_graph)

figure_file = here("results","figures","year_end_cases_graph.png")
ggsave(filename = figure_file, plot=year_end_cases_graph)
```
Correlation Analysis

This code chunk entails correlation testing through Spearman's and the creation of a matrix. First, lagged variables for case count was created for the data.
```{r}

data <- data %>%
  group_by(county) %>%
  arrange(date) %>%
  mutate(
    lag_1 = lag(new_cases, 1),
    lag_7 = lag(new_cases, 7),
    lag_14 = lag(new_cases, 14)
  ) %>%
  ungroup()

data <- data %>% drop_na()

# Create a data frame for matrix creation
cor_df <- data %>%
  select("New Cases" = new_cases, "1 Day Lag" = lag_1, "7 Day Lag" = lag_7, "14 Day Lag" = lag_14, "Population Density" = pop_density, "Retail Mobility" = retail_and_recreation_percent_change_from_baseline, "Workplace Mobility" = workplaces_percent_change_from_baseline, "Residential Mobility" = residential_percent_change_from_baseline, "Grocery Mobility" = grocery_and_pharmacy_percent_change_from_baseline,)

# Calculate correlations using Spearman's rank correlation because data is non parametric
cor_matrix <- cor(cor_df, method = "spearman")

# Melt for visualization
cor_matrix_melted <- melt(cor_matrix)

# Create plot for matrix
matrix <- ggplot(cor_matrix_melted, aes(x = Var1, y = Var2)) +
  geom_point(aes(size = abs(value), color = value)) +
  geom_text(aes(label = round(value, 2)), color = "black", size = 3) +
  scale_color_gradient2(low = "#BFA2DB", high = "#92C7CF", mid = "white",
                        midpoint = 0, limit = c(-1, 1), name = "Correlation") +
  scale_size(range = c(2, 10), guide = "none") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Correlation Heatmap (Spearman's)", x = "Variables", y = "Variables")

print(matrix)

figure_file = here("results","figures","matrix.png")
ggsave(filename = figure_file, plot=matrix)
```