### 2025-04-18 products/manuscript/Manuscript.qmd

---
title: "Assessing the Added Predictive Power of Google's Community Mobility Report for COVID-19 Forecasting on a County-Level"
author: "Vincent Nguyen"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/MADA.bib
csl: ../../assets/american-journal-of-epidemiology.csl
---

**Authors**

-   Vincent Nguyen$^{1}$

**Author affiliations**

1.  College of Public Health, University of Georgia, Athens, GA, USA.

$\land$ Corresponding author: vln27447\@uga.edu

{{< pagebreak >}}

# Summary/Abstract

*Write a summary of your project.*

{{< Examining the relationship between human mobility and infectious disease spread is essential for forecasting disease incidence and informing public health interventions. This analysis investigates community mobility patterns and their association with daily new COVID-19 case incidence in Georgia. Using county-level case counts and mobility trends from Google’s Community Mobility reports, machine learning models were constructed and evaluated; models included linear regression, Random Forest, and XGBoost. To assess the predictive power of mobility change metrics, two feature sets were utilized. The first set, referred to as “baseline”, included features known to affect disease incidence, such as lagged cases and population density. The second set included both the previous set’s features and metrics measuring changes in mobility (recorded as percent change from baseline). Models were trained and tuned using time-series blocked cross-validation and performance was evaluated R-squared, RMSE, and MAE metrics. Results indicated that models with mobility features could achieve modest improvements in RMSE and MAE, however, did not substantially increase R-squared, suggesting limited explanatory power. These findings suggest that while mobility data may slightly enhance short-term predictive accuracy, their overall predictive power in explaining COVID-19 incidence may be limited in this context. >}}

# 2 Introduction

## 2.1 COVID-19 Pandemic

The emergence of the novel corona-virus SARS-CoV-2 caused one of the most significant global health crises in modern history. First identified in Wuhan, China, in late 2019, the virus’s transmissibility and globalism led to its rapid spread worldwide, to leading the World Health Organization to declare it a pandemic on March 11, 2020. Just two days later, on March 13th, 2020, the Trump Administration declared a nationwide emergency due to the COVID-19 pandemic, marking the biggest virus outbreak since \_\_\_. In response, public health agencies implemented measures to curb the virus’ spread including travel restrictions, social distancing, and lockdown procedures (CDC 2024)[@centersfordiseasecontrolandprevention_2024_covid19].

## 2.2 Mobility Dynamics

Human mobility, a key driver[of respiratory disease transmi](https://www.nature.com/articles/s41598-024-64230-1)ssion, shifted significantly during the pandemic in response to public health policies and disease risk perception. Uniquely, the COVID-19 pandemic utilized a new form of physical distancing measures which was known as stay-at-home orders and colloquially, lockdowns. Lockdowns involved stringent stay-at-home orders, closure of non-essential businesses, and restrictions on public gatherings. Looking at the Wikipedia page shows a lack of prior history implementing lockdowns. Beyond these measures which lasted only a few weeks in Georgia, individual risk perception played into the compliance of other preventative behaviors (masking, social distancing, etc.). The[variability of individual risk perception](https://pmc.ncbi.nlm.nih.gov/articles/PMC9028425/#:~:text=The%20results%20indicate%20that%20a,%2C%20geographical%20factors%2C%20and%20timing.)has led to complex mobility patterns during the pandemic, for example, surges in mobility amidst large case outbreaks. In order to analyze and quantify this relationship, machine learning models were adapted.

\[insert stats about pandemic\]

## 2.3 Machine Learning

[Machine Learning](https://pmc.ncbi.nlm.nih.gov/articles/PMC9949554/) refers to a class of data-driven algorithms that aim to analyze associations and relationships found in data. These techniques can be supervised or unsupervised; in supervised ML, models are given labeled inputs/features to derive and predict an output. These models aim to also approximate the relationship between outputs and inputs, quantify predictions, or approximate classification tasks. Beyond regular statistical analysis, machine learning methods offer powerful tools for forecasting infectious disease trends by identifying complex, nonlinear relationships between predictors and outcomes. S[ome common m](https://pmc.ncbi.nlm.nih.gov/articles/PMC8219638/)odels in disease forecasting include ARMA, ARIMA, LASSO, XGBoost, and various neural network techniques. Several autho[rs](https://pmc.ncbi.nlm.nih.gov/articles/PMC8731233/)have utilized case counts, estimations, demographics, and more to forecast disease trends. However, the predictive value of real-time mobility data, particularly in the context of an evolving pandemic, has not been assessed through modeling. This study aims to explore whether mobility dynamics, form Google’s Community Mobility Reports, can enhance predictive performance when modeling COVID-19 transmission at the county level.

# 3 Data

## 3.1 Google’s Community Mobility Report

During the pandemic, Google began to collect aggregated, anonymized data from users utilizing Google products (apps, phones, etc.) to track changes in mobility. Within the data, Google measures mobility as a percent change difference from baseline measurements; for example, a -45% change in retail and recreation mobility indicates a 45% reduction in movement to those categorical locations. These measurements are stratified by county, however, with the implication of technology use, are limited to counties with enough Google users. Additionally, two mobility metrics, transit stations and recreational parks, were omitted in this analysis due to incompleteness.

## 3.2 John Hopkins University COVID-19 Case Data

The COVID-19 Data Repository by the Center for Systems Science and Engineering at Johns Hopkins University is a comprehensive data set that tracks global COVID-19 cases, recoveries, and deaths. This data was recorded daily for several years across every county in the US.

## 3.3 Final Data-set

After data wrangling and cleaning for completeness, 44 counties were included in the analysis. An 80% completeness of data per county was required for inclusion into analysis, resulting in the removal of 115 counties. Many of Georgia’s counties are rural and as such, lacked many mobility metrics. The data was filtered from March 14t 2020 to March 14 2022. This was selected specifically as March 14th marks the date of Governor Brian Kemp’s announcement of Georgia's Public Health State of Emergency.

{{< pagebreak >}}

# 4 Methods

## 4.1 Model Selection

As informed by Alfred and Obit 2021, machine learning models have wide application in forecasting outbreaks and disease incidence. In their review, they outline several common applications of regression and classification models. Their review, along with class content and data structure, informed model selection. Previous iterations of the analysis indicated that the data was non-normal and non-stationary which violated assumptions in common time series models like ARIMA. Linear regression was chosen as a standard model while Random Forest and XGBoost were chosen for their lack of assumptions required for analysis.

### Linear Regression

A statistical modeling method used to quantify the relationship between a continuous outcome variable and one or more predictors. In its simplest form, linear regression estimates, or fits, a linear equation explaining the relationship by minimizing the sum of squared differences between residuals. Linear regression assumes

### Random Forest

(https://www.stat.berkeley.edu/\~breiman/randomforest2001.pdf)\

A ensemble machine learning method that builds multiple decision trees and aggregates their predictions to improve accuracy and reduce overfitting. A decision tree is a modeling technique that partitions  data into subsets based on the input features; it aims to minimize variance of the outcome variable in each subgroup. These splits form a tree-like structure whe -. Decision trees are powerful and can capture nonlinear complex relationships, however, can overfit training data.  Random forests are an average of these trees to produce a final prediction and can produce more stable/accurate models as a result.

### XGBoost

(https://arxiv.org/abs/1603.02754): Also known as Extreme Gradient Boosting, XGBoost is a machine learning algorithm based on decision tress, however, instead of averaging trees, it tries to correct errors made by previous ones. This technique is known as boosting. In simple terms, XGBoost makes predictions with a single tree and looks at the errors and creates a second tree focused on minimizing the errors.

Two sets of predictors were chosen to assess the effect mobility changes have on case incidence. The first set, the baseline predictors, included population density and 3 spaced out lagged variables for case counts (1 day, 7 days, and 14 days). These were chosen because of  their [known influence](https://pmc.ncbi.nlm.nih.gov/articles/PMC8983058/#:~:text=This%20study%20has%20shown%20that,data%20for%20COVID%E2%80%9019%20cases.) on disease transmission. The second set, the mobility predictors, included population density, the 3 spaced out lagged variables, and mobility indicators from Google’s report (e.g., workplace, grocery, retail, transit).

## 4.2 Cross Validation

To evaluate model performance while accounting for the temporal structure of the data, a time-blocked cross-validation strategy was employed. The dataset was partitioned into eight non-overlapping temporal folds, each representing a three-month interval spanning March 2020 to March 2022. Each observation (a day for a corresponding county) was assigned a fold based on its calendar month. Grouped k-fold cross-validation using these time-based identifiers were used to ensure data from the future could not leak into the training set.

## 4.3 Model Evaluation and Metrics

### R-squared

A statistical measure that explains the proportion of the data variation that can be explained by the model. A higher R-squared value generally is indicative of a better-fitting model, although could be misleading when over fitting occurs.

### RMSE

Root Mean Square Deviation is a statistical measure that averages the magnitude of the errors in the model’s predictions.

Arithmetically, it is the square root of the average squared differences between the predicted and actual values. A lower RMSE indicates a better predictive performance.

### MAE

Mean Absolute Error is a statistical measure that also measures prediction error, however, utilizes the absolute difference between predicted and actual values instead of the squaring the errors. It is calculated by taking the average absolute difference between each predicted and actual value. Because of its computation, MAE is less sensitive to large errors, like outliers, than RMSE. MAE may be a more reliable metric for this data set due to the presence of large outbreaks that can inflate the errors.

Linear regression, Random Forest, and XGBoost models were trained using the baseline set of features to predict COVID-19 case incidence. These baseline models served as reference points to evaluate the added predictive value of mobility-based features and to assess the potential association between changes in population mobility and COVID-19 case incidence.

After the initial RF and XGBoost models were made, their hyperparameters were tuned to improve predictive performance.

## 4.4 Software

The analysis was conducted on RStudio 4.3.2 on Windows 11. The following R packages were used: [ggplot2](https://cran.r-project.org/web/packages/ggplot2/index.html), [broom](https://cran.r-project.org/web/packages/broom/index.html), [here](https://cran.r-project.org/web/packages/here/index.html), [glmnet](https://cran.r-project.org/web/packages/glmnet/index.html), [MASS](https://cran.r-project.org/web/packages/MASS/index.html), [tidymodels](https://cran.r-project.org/web/packages/tidymodels/index.html), [dplyr](https://cran.r-project.org/web/packages/dplyr/index.html), [rsample](https://cran.r-project.org/web/packages/rsample/index.html), [parsnip](https://cran.r-project.org/web/packages/parsnip/index.html), [future](https://cran.r-project.org/web/packages/future/index.html), [vip](https://cran.r-project.org/web/packages/vip/index.html), [zoo](https://cran.r-project.org/web/packages/zoo/index.html), [patchwork](https://cran.r-project.org/web/packages/patchwork/index.html), [e1071](https://cran.r-project.org/web/packages/e1071/index.html), [scales](https://cran.r-project.org/web/packages/scales/index.html), [RColorBrewer](https://cran.r-project.org/web/packages/RColorBrewer/index.html), [corrplot](https://cran.r-project.org/web/packages/corrplot/index.html), [reshape2](https://cran.r-project.org/web/packages/reshape2/index.html), [tidyr](https://cran.r-project.org/web/packages/tidyr/index.html), [lubridate](https://cran.r-project.org/web/packages/lubridate/index.html), [readxl](https://cran.r-project.org/web/packages/readxl/index.html), [stringr](https://cran.r-project.org/web/packages/stringr/index.html), [skimr](https://cran.r-project.org/web/packages/skimr/index.html), [tigris](https://cran.r-project.org/web/packages/tigris/index.html), and [sf](https://cran.r-project.org/web/packages/sf/index.html).

{{< pagebreak >}}

# 5 Results

## 5.1 Exploratory/Descriptive analysis

```{r}
#| label: cases_graph
#| fig-cap: "Total COVID-19 Cases in Georgia over time"
#| echo: FALSE
knitr::include_graphics(here("results","figures","cases_graph.png"))
```

```{r}
#| label: incidenceplot
#| fig-cap: "Daily New COVID-19 Cases overtime in the GA (2020-2022)"
#| echo: FALSE
knitr::include_graphics(here("results","figures","incidence_graph.png"))
```

```{r}
#| label: mobiltiygraph
#| fig-cap: "Percent Changes in Mobility per week in GA"
#| echo: FALSE
knitr::include_graphics(here("results","figures","mobilitygraph.png"))
```

```{r}
#| label: mobilitygrid
#| fig-cap: "Percent Changes in Mobility per week in GA"
#| echo: FALSE
knitr::include_graphics(here("results","figures","mobility_grid.png"))
```

```{r}
#| label: mobilitymap
#| fig-cap: "Average of mobility changes on 10/15/2022 mapped to the U.S."
#| echo: FALSE
knitr::include_graphics(here("results","figures","mobilitymap.png"))
```

## Correlation Analysis

```{r}
#| label: correlation
#| fig-cap: "Correlation Matrix of Predictors using Spearman's"
#| echo: FALSE
knitr::include_graphics(here("results","figures","matrix.png"))
```
 In addition to visual analysis, correlation analysis was done. The non-parametric method, Spearman’s rank correlation coefficient, was used to assess correlation among predictors. Analysis reveals that lagged case counts were strongly correlated with daily incidence while in contrast mobility variables demonstrated weak relationships. This is suggestive of lagged variables's strong predictive power and mobility metric’s weak power.

## Full analysis

*Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here.*

{{< pagebreak >}}

# Discussion

## Summary and Interpretation

*Summarize what you did, what you found and what it means.*

## Strengths and Limitations

*Discuss what you perceive as strengths and limitations of your analysis.*

## Conclusions

*What are the main take-home messages?*

*Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end*

This paper [@leek2015] discusses types of analyses.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like.

{{< pagebreak >}}

# References