### 2025-04-21 code/processing-code/processing-file.qmd
Meta data

---
title: "Cleaning Script"
author: "Vincent Nguyen"
date: "2025-02-21"
output: html_document
---

# Setup

Load needed packages. Make sure they are installed.

```{r}
# load needed packages
library(here) 
library(dplyr)
library(tidyr)
library(lubridate)
library(readxl)
library(stringr)
library(skimr)

library(tigris)
library(sf)
```

# Data loading

```{r}
#Google data is in three parts; load and join
mobility_data_2020 <- read.csv(here("data", "raw-data", "2020_US_Region_Mobility_Report.csv")) %>%
  filter(sub_region_1 == "Georgia")
mobility_data_2021 <- read.csv(here("data", "raw-data", "2021_US_Region_Mobility_Report.csv"))%>%
  filter(sub_region_1 == "Georgia")
mobility_data_2022 <- read.csv(here("data", "raw-data", "2022_US_Region_Mobility_Report.csv"))%>%
  filter(sub_region_1 == "Georgia")

# Combine columns and remove irrelevant rows
combined_mobility_data <- bind_rows(mobility_data_2020, mobility_data_2021, mobility_data_2022)

# jhk covid time series cases
time_series_covid <- read.csv(here("data", "raw-data", "time_series_covid19_confirmed_US.csv")) %>%
  filter(Province_State == "Georgia")

# county population information
ch_pop_data <- read_excel(here::here("data", "raw-data", "co-est2023-pop.xlsx"), skip = 3)

```

# Cleaning and Wrangling

For this cleaning file, the workflow focused on preparing the data frames for joining and cleaning out columns as joins and mutations occurred.\

## County Population Information

The county population count data was first. Unnecessary rows were identified through the Rstudio view() feature and removed. Also, the data was converted into a long format for easy joining and analysis later.

```{r}
# Remove the second column
ch_pop_data <- ch_pop_data[, -2]

# Convert to a long format
ch_pop_data <- ch_pop_data %>%
  pivot_longer(
    cols = c(`2020`,`2021`, `2022`, `2023`),    
    names_to = "year",        
    values_to = "population_count"              
  )

# Fix the first column
ch_pop_data <- ch_pop_data %>%
  mutate(
    county = str_remove(`...1`, "^\\."),            
    county = str_remove(county, "County"),     
    county = str_to_lower(county),                  
    state = str_extract(county, "(?<=,).*"),        
    county = str_remove(county, ",.*")              
  ) %>%
  mutate(
    state = str_to_lower(str_trim(state))           
  )

# Change year column to be double for joining later
ch_pop_data$year <- as.double(ch_pop_data$year)

# Clean names to prepare for joining
ch_pop_data <- ch_pop_data %>%
  mutate(
    county = str_trim(str_to_lower(county)),  
    state = str_trim(str_to_lower(state))     
  )

# Only include Georgia counties
ch_pop_data <- ch_pop_data %>%
  filter(state == "georgia")

```

## Google Community Mobility Reports

Next, upon visual inspection in preparation for joining, it became apparent that the mobility report had many missing values. To quantify this, NA values were counted across columns and skim() was used.

After noticing a lack of data in the transit and parks category, they were removed. This does remove some important info, however, the completeness of the data was priortized instead. Additionally, not all incomplete rows were removed so completeness can be assessed later.

The data frame was also cleaned and prepared for joining.

```{r}
# Get a glimpse at what data is complete
colSums(is.na(combined_mobility_data))

skim(combined_mobility_data)

# Ok it is clear that some locations are absent from GA as a whole. Based on my own knowledge and this data, I should consider removing transit stations as public transit is not common in GA. Additionally, interestingly, parks are absent. Perhaps Google defines parks differently? For now, I will remove the category of mobility. I deal with completeness later in the code.
combined_mobility_data <- combined_mobility_data %>% 
  dplyr::select(-transit_stations_percent_change_from_baseline, -parks_percent_change_from_baseline)

# Prepare mobility report data for joining
mobility_df <- combined_mobility_data %>%
  rename(county = sub_region_2, state = sub_region_1) %>%
  mutate(county = tolower(county), 
         state = tolower(state),
         county = gsub(" county$", "", county),  # Remove "county" from county name
         date = as.Date(date, format = "%Y-%m-%d"))  # Correct format for "YYYY-MM-DD"
```

## John Hopkins Case Count Time Series

Lastly, the JHU data was prepared. After inspection, it is clear that the Time Series data is wide while the Mobility Report is long. For easier analysis and joining, the Time Series data was converted into a long format using pivot_longer().

```{r}

# Pivot the time series data and change names around
cases_long <- time_series_covid %>%
  pivot_longer(
    cols = starts_with("X"),  
    names_to = "date",
    values_to = "total_cases"
  ) %>%
  mutate(date = gsub("^X", "", date),  # Remove leading x for dates to join data sets in a bit
         date = gsub("\\.", "/", date),  # Replace dots with slashes
         date = as.Date(date, format = "%m/%d/%y"))  # Use "%y" instead of "%Y"

# Standardize county & state names (convert to lowercase to avoid mismatches)
cases_long <- cases_long %>%
  rename(county = Admin2, state = Province_State) %>%
  mutate(county = tolower(county), state = tolower(state))
```

## Filtering and Joining

After preparing the data sets and visual inspection, it was clear the dates for the data sets differed. The view() feature in Rstudio was used to find the overlap.

Once filtered, all three data sets were joined, an incidence per 10k column was added, and more column dropping occurred.

```{r}
# Filter cases_long and mobility_df based on the date range
cases_long_filtered <- cases_long %>%
  filter(date >= "2020-02-15" & date <= "2022-10-15")

mobility_df_filtered <- mobility_df %>%
  filter(date >= "2020-02-15" & date <= "2022-10-15")

# Merge the data sets together
merged_df <- cases_long_filtered %>%
  inner_join(mobility_df_filtered, by = c("county", "state", "date"))

# Additionally, add a column about new cases for further analysis down the line.
merged_df<- merged_df %>%
  arrange(state, county, date) %>%  
  group_by(state, county) %>%  
  mutate(new_cases = c(NA, diff(total_cases))) %>%  
  ungroup()

# Join the county population count data
merged_df<- merged_df %>%
  mutate(year = year(date)) %>%
  left_join(
    ch_pop_data %>% 
      dplyr::select(year, population_count, county, state),
    by = c("state" = "state", "county" = "county", "year" = "year")
  )

# Create column for incidence rate per 10,000 (10,000 because numbers tend to be small when per 100k)
merged_df <- merged_df %>%
  mutate(incidence_rate = (new_cases / population_count) * 10000)

# Clean up and Removal of unnecessary columns
merged_df <- merged_df %>% dplyr::select(-UID, -iso2, -iso3, -code3, -country_region_code, -country_region, -iso_3166_2_code, -place_id, -census_fips_code, -metro_area, -FIPS, -Lat, -Long_, -Combined_Key)
```

## Implementation of SF Package for Population Density Calculation

To help explain transmission patterns, population density was calculated using the shape files found in the package, sf.

```{r}
library(tigris)
library(sf)

# Load county shapefile for GA
ga_counties <- counties(state = "GA", cb = TRUE, class = "sf")

# Make county column lowercase for merging
ga_counties <- ga_counties %>%
  mutate(NAME = tolower(NAME))

# Merge with processed population data (assumed already cleaned)
ga_counties <- ga_counties %>%
  left_join(ch_pop_data, by = c("NAME" = "county"))

# Calculate county area in square kilometers
ga_counties <- ga_counties %>%
  mutate(area_km2 = as.numeric(st_area(geometry)) / 1e6)  # Convert m² to km²

# Compute population density (people per km²) and remove unnecessary data
ga_counties <- ga_counties %>%
  mutate(pop_density = population_count / area_km2) %>%
  dplyr::select(pop_density, year, NAME)

# Add in population density column and remove the geometry from the df
merged_df <- merged_df %>%
  left_join(ga_counties, by = c("county" = "NAME", "year" = "year")) %>%
  as_tibble() %>%
  dplyr::select(-geometry)

```

## Additional Cleaning and Filtering

This code chunk focuses on cleaning some of the values to not be NA and some final filtering.

```{r}
# Clean negative and NA values
merged_df$new_cases[merged_df$new_cases < 0] <- 0
merged_df$new_cases[is.na(merged_df$new_cases)] <- 0
merged_df$incidence_rate[is.na(merged_df$incidence_rate)] <- 0

# Filter for 2 years (outside of this range generally is 0)
last_case_date <- as.Date("2022-03-14")
first_case_date <- as.Date("2020-03-14")

merged_df <- merged_df %>%
  filter(date >= first_case_date) %>%
  filter(date <= last_case_date) %>%
  group_by(county)

# Convert new_cases to integer
merged_df<- merged_df %>% 
  mutate(new_cases = as.integer(new_cases))

```

## Evaluation and Filtering for Completeness of Data

As a final step, missing values in the mobility data were addressed by evaluating the completeness of each county's observations; the focus was placed on identifying days in which all mobility metrics were reported. Counties with a high proportion of incomplete records were excluded, as these gaps could mask meaningful trends and introduce bias. To mitigate this issue, counties with less than 80% data completeness were removed to support more robust and reliable analyses.

```{r}
# check the completeness by county
completeness_by_county <- merged_df %>%
  dplyr::select(retail_and_recreation_percent_change_from_baseline,  
                grocery_and_pharmacy_percent_change_from_baseline,
                workplaces_percent_change_from_baseline,
                residential_percent_change_from_baseline) %>%
  group_by(county) %>%
  rowwise() %>%
  mutate(row_complete = all(!is.na(c_across(everything())))) %>%
  ungroup() %>%
  group_by(county) %>%
  summarise(
    total_rows = n(),
    complete_rows = sum(row_complete),  # Count of rows that are complete
    completeness_rate = complete_rows / total_rows  # Proportion of complete rows
  ) %>%
  arrange(desc(completeness_rate))

# Set the completeness threshold of 80%
completeness_threshold <- 0.80

# Filter counties based on the completeness rate
counties_to_include <- completeness_by_county %>%
  filter(completeness_rate >= completeness_threshold) %>%
  pull(county)

# Filter the dataset to include only counties with at least 80% completeness
merged_df_clean <- merged_df %>%
  filter(county %in% counties_to_include)
```

The cleaning and wrangling process is complete.

To summarize, the three data sets were joined, cleaned, filtered, and mutated to create columns (population density and incidence) for analysis later.

# Save data

```{r}
save_data_location <- here::here("data","processed-data","processeddata.rds")
saveRDS(merged_df_clean, file = save_data_location)

# Saved a more complete case data set for visualizations
save_data_location <- here::here("data","processed-data","processedcasedata.rds")
saveRDS(cases_long_filtered, file = save_data_location)
```