### 2025-04-21 code/eda-code/eda.qmd

---
title: "Exploratory Data Analysis"
date: "2025-02-21"
output: html_document
---

This Quarto file loads the cleaned data and does some exploring. \# Setup

```{r}
# load needed packages
library(here) 
library(dplyr)
library(ggplot2)
library(zoo)
library(patchwork)
library(e1071)
library(scales)
library(RColorBrewer)
library(corrplot)
library(reshape2)
library(lubridate)
library(tidyr)
```

Load the data from processingfile-v1 file.

```{r}

# processed data
data <- readRDS(here("data", "processed-data", "processeddata.rds"))

# Load full case data
case_data <- readRDS(here("data", "processed-data", "processedcasedata.rds"))

# Make incidence rate non negative
data<- data %>%
   mutate(incidence_rate = if_else(incidence_rate < 0, 0, incidence_rate))
```

Data Visualization

Since new case incidence is the variable of interest, it was visually inspected to understand the distribution.

From the graphs, kurtosis test, and skewness test, it is apparent that daily new case incidence is not normally distribued and heavily skewed towards 0 and other lower values. The second graph, new cases over 500, was graphed to show how certain days do reach high case incidence.
```{r}
# Plot density of case incidence
density_cases <- ggplot(data, aes(x=new_cases)) +
    geom_density(fill="#69b3a2", color="#69b3a2", alpha=0.8) +
  labs(title = "Distribution of Daily New Case Incidence", 
       x = "Daily New Cases",
       y = "Density") +
  theme_bw()

# print
print(density_cases)

# Save
figure_file = here("results","figures","density_cases.png")
ggsave(filename = figure_file, plot=density_cases)

# plot density of case incidence over 500
density_cases_500 <- data %>%
  filter(new_cases > 500) %>%
  ggplot( aes(x=new_cases)) +
    geom_density(fill="#69b3a2", color="#69b3a2", alpha=0.8) +
  labs(title = "Distribution of Daily New Case Incidence (Over 500)", 
       x = "Daily New Cases",
       y = "Density") +
  theme_bw()

# print
print(density_cases_500)

# save
figure_file = here("results","figures","density_cases_500.png")
ggsave(filename = figure_file, plot=density_cases_500)

print(kurtosis(data$new_cases))
print(skewness(data$new_cases))
```

A histogram of population density was created. I hypothesize that population density plays a large role in case incidence. Population density is also skewed, however, there are some notable counties that do have large populationd densities.
```{r}
# average population density throughout all three years
data_summary <- data %>%
  group_by(county) %>%
  summarize(pop_density = mean(pop_density, na.rm = TRUE))

# plot summary of population density
histo_pop_density <- ggplot(data_summary, aes(x = pop_density)) + 
  geom_histogram(bins = 30, fill = "#69b3a2", color = "#69b3a2", alpha = 0.7) +
  labs(title = "Histogram of Population Density (2-Year Average per County)", 
       x = "Population Density (Calculated as Population Count / km squared)", y = "Frequency") + theme_bw()

# print
print(histo_pop_density)

# save
figure_file = here("results","figures","histo_pop_density.png")
ggsave(filename = figure_file, plot=histo_pop_density)
```

A 7-day rolling average was applied to the mobility data as the original iteration was too jagged. This code entails the creation of a data frame for 7 day averages.

```{r}

# Create dataframe of average mobility per county per date
ga_average_mobility <- data %>%
  group_by(date) %>%
  summarise(
    retail_mean = mean(retail_and_recreation_percent_change_from_baseline, na.rm = TRUE),
    grocery_mean = mean(grocery_and_pharmacy_percent_change_from_baseline, na.rm = TRUE),
    workplaces_mean = mean(workplaces_percent_change_from_baseline, na.rm = TRUE),
    residential_mean = mean(residential_percent_change_from_baseline, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  arrange(date) %>%
  mutate(
    retail_mean_7day = rollmean(retail_mean, k = 7, fill = NA, align = "right"),
    grocery_mean_7day = rollmean(grocery_mean, k = 7, fill = NA, align = "right"),
    workplaces_mean_7day = rollmean(workplaces_mean, k = 7, fill = NA, align = "right"),
    residential_mean_7day = rollmean(residential_mean, k = 7, fill = NA, align = "right")
  )

# Set custom colors
custom_colors <- c(
  "Grocery & Pharmacy" = "#69b3a2",
  "Workplaces" = "#BFA2DB",
  "Residential" = "#F2C29B",
  "Retail" = "#DBA2A2"
)

# Plot the 7-day averages for each mobility type
mobility_graph <- ggplot(ga_average_mobility, aes(x = date)) +
  geom_line(aes(y = grocery_mean_7day, color = "Grocery & Pharmacy")) +
  geom_line(aes(y = workplaces_mean_7day, color = "Workplaces")) +
  geom_line(aes(y = residential_mean_7day, color = "Residential")) +
  geom_line(aes(y = retail_mean_7day, color = "Retail")) +
  scale_color_manual(values = custom_colors) +
  labs(
    title = "7-Day Moving Average of Percent Change in Mobility in Georgia Over Time",
    x = "Date",
    y = "7-Day Average of Percent Change in Mobility",
    color = "Mobility Type"
  ) +
  theme_bw()

# print
print(mobility_graph)

# save graph
figure_file = here("results","figures","mobilitygraph.png")
ggsave(filename = figure_file, plot=mobility_graph)
```

For better legiiblity, all 4 mobiltiy types were graphed separately.

```{r}
# Create one plot per mobility type
grocery <- ggplot(ga_average_mobility, aes(x = date, y = grocery_mean_7day)) +
  geom_line(color = "#69b3a2") +
  labs(title = "Grocery & Pharmacy", x = "Date", y = "7-Day Avg % Change") +
  theme_bw() +
  coord_cartesian(ylim = c(-50, 20))

workplace <- ggplot(ga_average_mobility, aes(x = date, y = workplaces_mean_7day)) +
  geom_line(color = "#BFA2DB") +
  labs(title = "Workplaces", x = "Date", y = "7-Day Avg % Change") +
  theme_bw() +
  coord_cartesian(ylim = c(-50, 20))

resident <- ggplot(ga_average_mobility, aes(x = date, y = residential_mean_7day)) +
  geom_line(color = "#F2C29B") +
  labs(title = "Residential", x = "Date", y = "7-Day Avg % Change") +
  theme_bw() +
  coord_cartesian(ylim = c(-50, 20))

retail <- ggplot(ga_average_mobility, aes(x = date, y = retail_mean_7day)) +
  geom_line(color = "#DBA2A2") +
  labs(title = "Retail", x = "Date", y = "7-Day Avg % Change") +
  theme_bw() +
  coord_cartesian(ylim = c(-50, 20))

# Combine in 2x2
mobility_grid <- (grocery | workplace) / (resident | retail)

# print for visual inspection
print(mobility_grid)

# Save it
figure_file = here("results", "figures", "mobility_grid.png")
ggsave(filename = figure_file, plot = mobility_grid)

```

After visually inspecting the new cases over time graph, heteroscedastity and non-stationairty are of concern. Large outbreaks can skew these metrics and even when tested, could be misleading. Additionally, assumptions of homoscedastity are required for some tests of stationairty and some assumptions of normality are required for some tests of homoscdastity. Because of these concerns, I chose to focus on models like XGBoost and Random Forest which do not require these assumptions.

```{r}
library(scales)

# Get first and last date of data
first_date <- min(data$date)
last_date <- max(data$date)

# Create graph of total case count through the 3 years
cases_graph <- case_data %>%
  group_by(date) %>%
  summarise(total_cases = sum(total_cases, na.rm = TRUE)) %>%
  arrange(date) %>%
  ggplot(aes(x = date, y = total_cases)) +
    geom_area(fill = "#69b3a2", alpha = 0.5) +
    geom_line(color = "#69b3a2") +
    ylab("Total Cases (in GA)") +
    scale_y_continuous(labels = label_number()) +
   scale_x_date(
      breaks = seq(from = first_date, to = last_date, by = "6 months"),
      labels = date_format("%b %Y")
    ) +
    theme_bw() +
  labs(title = "Total Cases (in GA) Over Time",
       x = "Date (2020 - 2022)",
       y = "Total Cases")

# print
print(cases_graph)

# save
figure_file = here("results","figures","cases_graph.png")
ggsave(filename = figure_file, plot=cases_graph)

# plot incidence
incidence_graph <- data %>%
  group_by(date) %>%
  summarise(new_cases = sum(new_cases, na.rm = TRUE)) %>%
  arrange(date) %>%
  ggplot(aes(x = date, y = new_cases)) +
    geom_area(fill = "#69b3a2", alpha = 0.5) +
    geom_line(color = "#69b3a2") +
    ylab("Daily New Cases") +
     scale_x_date(
      date_breaks = "3 months",
      date_labels = "%b %Y"
    ) +
    theme_bw() +
  labs(title = "Daily New Cases (in GA) Over Time",
       x = "Date (2020 - 2022)",
       y = "Total Cases")

# print
print(incidence_graph)

# save
figure_file = here("results","figures","incidence_graph.png")
ggsave(filename = figure_file, plot=incidence_graph)
```

This code chunk plots a scatter plot of a county's population count against its final total cases. 
```{r}
# create df with total case count at the end of the year
year_end_cases <- data %>%
  mutate(year = year(date)) %>%
  group_by(county, year) %>%
  slice_max(order_by = date, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(county, year, total_cases)

# pick population density of the years
pop_density_info <- data %>%
  mutate(year = year(date)) %>%
  select(county, year, pop_density) %>%
  distinct()

# combine dfs for ploting
plot_df <- left_join(year_end_cases, pop_density_info, by = c("county", "year"))

# plot
year_end_cases_graph <- ggplot(plot_df, aes(x = pop_density, y = total_cases)) +
  geom_point(color = "#69b3a2", alpha = 0.6) +
  facet_wrap(~year) +
  labs(
    x = "Population Density",
    y = "Total COVID Cases (End of Year)",
    title = "Year-End COVID Cases vs. Population Density by County"
  ) +
  theme_bw()

# print
print(year_end_cases_graph)

# save
figure_file = here("results","figures","year_end_cases_graph.png")
ggsave(filename = figure_file, plot=year_end_cases_graph)
```
Correlation Analysis

This code chunk entails correlation testing through Spearman's and the creation of a matrix. First, lagged variables for case count was created for the data.
```{r}
# create lagged variables
data <- data %>%
  group_by(county) %>%
  arrange(date) %>%
  mutate(
    lag_1 = lag(new_cases, 1),
    lag_7 = lag(new_cases, 7),
    lag_14 = lag(new_cases, 14)
  ) %>%
  ungroup()

# Drop NAs made as a result
data <- data %>% drop_na()

# Create a data frame for matrix creation
cor_df <- data %>%
  select("New Cases" = new_cases, "1 Day Lag" = lag_1, "7 Day Lag" = lag_7, "14 Day Lag" = lag_14, "Population Density" = pop_density, "Retail Mobility" = retail_and_recreation_percent_change_from_baseline, "Workplace Mobility" = workplaces_percent_change_from_baseline, "Residential Mobility" = residential_percent_change_from_baseline, "Grocery Mobility" = grocery_and_pharmacy_percent_change_from_baseline,)

# Calculate correlations using Spearman's rank correlation because data is non parametric
cor_matrix <- cor(cor_df, method = "spearman")

# Melt for visualization
cor_matrix_melted <- melt(cor_matrix)

# Create plot for matrix
matrix <- ggplot(cor_matrix_melted, aes(x = Var1, y = Var2)) +
  geom_point(aes(size = abs(value), color = value)) +
  geom_text(aes(label = round(value, 2)), color = "black", size = 3) +
  scale_color_gradient2(low = "#BFA2DB", high = "#69b3a2", mid = "white",
                        midpoint = 0, limit = c(-1, 1), name = "Correlation") +
  scale_size(range = c(2, 10), guide = "none") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Correlation Heatmap (Spearman's)", x = "Variables", y = "Variables")

# Print
print(matrix)

# save
figure_file = here("results","figures","matrix.png")
ggsave(filename = figure_file, plot=matrix)
```